---
title: "Statistical Inference Course Project Part 2"
author: "jacethedatascientist"
date: "August 12, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statistical Inference Course Project Part 2
#### *Jace Galleon*

***

## I. Overview

This is the second part of the Statistical Inference Course Project from Coursera.

In this project, we're going to analyze the ToothGrowth data in the R datasets package.

***

## II. Objectives

1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4. State your conclusions and the assumptions needed for your conclusions.

***

## III. Data Processing
### A. Exploration
```{r, cache = TRUE, echo = TRUE}
data("ToothGrowth")
print(names(ToothGrowth))
print(summary(ToothGrowth))
print(paste("Number of Rows: ", nrow(ToothGrowth)))
print(sapply(ToothGrowth, class))
```

### B. Preprocessing
1. Since the dosage is a level not an actual value, it will be converted to factor.
```{r, cache = TRUE, echo = TRUE}
procData <- ToothGrowth
procData$dose <- as.factor(procData$dose)
```
2. Relationship between the Tooth Length and Supplements
```{r, cache = TRUE, echo = FALSE}
meanSupp <- split(procData$len, procData$supp)
sapply(meanSupp, mean)
```
3. Relationship between the Tooth Length and Dosage Level
```{r, cache = TRUE, echo = FALSE}
meanDose <- split(procData$len, procData$dose)
sapply(meanDose, mean)
```
4. Variance by Dosage Level
```{r, cache = TRUE, echo = TRUE}
sapply(meanDose, var)
```
5. Variance by Supplements
```{r, cache = TRUE, echo = TRUE}
sapply(meanSupp, var)
```

***

\newpage

## IV. Analysis
### Tooth Growth Comparison
#### A. By Supplements
##### Hypotheses Testing
**Null** : *The supplements has no impact on Tooth Growth*  
**Alternative** : *At least one of the supplements has impact on Tooth Growth*

The P-value is **0.06063** and is **greater than the 0.05 significance value**, then, *we don't have any sufficient data to reject the null hypothesis*.

Thus, both the supplements **has no impact** on Tooth Growth

#### B. By Dosage Level
##### Hypotheses Testing
**Null** : *Higher dosage level has no impact on Tooth Growth*  
**Alternative** : *Higher dosage level has impact on Tooth Growth*

The P-value is **almost 0** and obviously **lesser than the 0.05 significance value**, then, *we have sufficient data to reject the null hypothesis*.

Thus, higher dosage level **has impact** on Tooth Growth.

***

## V. Conclusion
### Summary
#### By using the T-Test, we can now say that:
1. The supplements (Orange Juice and Vitamin C) **has no impact** on the Tooth Growth.
2. Higher dosage level (0.2 or higher) **has impact** on Tooth Growth

### Assumptions
* each subject is randomly assigned
* each subject is a representative of the population
* observations are independent

***

*This formally ends the Part 2 of the Course Project. Thank You!*

\newpage

## VI. Additional Chart

### ToothGrowth Scatter Plot
```{r, cache = TRUE, echo = FALSE}
plot(ToothGrowth)
```

### ToothGrowth Box Plot
```{r, cache = TRUE, echo = FALSE}
library(ggplot2)
ggplot(aes(x = supp, y = len), data = procData) + geom_boxplot(aes(fill = supp)) + xlab("Application Method") + ylab("Tooth Length [mm]") + ggtitle("Tooth Length by Supplements") + scale_fill_manual(name = "Application method", values = c("darkorange1","darkgoldenrod1"), labels = c("Orange juice (OJ)", "Vitamin C (VC)"))
ggplot(aes(x = dose, y = len), data = procData) + geom_boxplot(aes(fill = dose)) + xlab("Dosage Level") + ylab("Tooth Length [mm]") + ggtitle("Tooth Length by Dosage Level") + scale_fill_manual(name = "Dosage Level", values = c("darkturquoise","cornflowerblue","darkcyan"), labels = c("0.5mg", "1mg", "2mg"))
```

### T-Test By Supplements
```{r, cache = TRUE, echo = FALSE}
t.test(procData$len[procData$supp == "OJ"], procData$len[procData$supp == "VC"], paired = FALSE, var.equal = FALSE)
```

### T-Test By Dosage Level
```{r, cache = TRUE, echo = FALSE}
t.test(procData$len[procData$dose == 2], procData$len[procData$dose != 2], paired = FALSE, var.equal = TRUE)
```